```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1
1. Find all flights that:
a. Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```
b. Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```
c. Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA", "DA"))
```
d. Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% c(7, 8, 9))
```
e. Arrived more than two hours late, but didn't leave late
```{r}
filter(flights, arr_delay > 120, dep_delay < 1)
```
f. Were delayed by at least an hour, but made up over 30
minutes in flight
```{r}
filter(flights, dep_delay >= 60, air_time >= 30)
```
g. Departed between midnight and 6 a.m. (inclusive)
```{r}
filter(flights, dep_time == 2400 | dep_time <= 600)
```

##  Exercise 2
2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

between() checks if the first argument is greater than/equal to the second argument and less than/equal to the third argument. This can simplify code that used multiple inequalities or where several continuous possible values of a variable were listed out individually. For example, question 1.d can be simplified to:
```{r}
filter(flights, between(month, 7, 9))
```

##  Exercise 3
3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
By doing:
```{r}
summary(flights)
```
We can tell that there are 8255 flights with an unknown dep_time. Other missing variables include 8255 missing values for dep_delay, 8713 for arr_time, and 9430 for arr_delay and air_time. These could represent flights where data was not recorded or lost, flights that were rerouted, or flights that never arrived/departed.

# Chapter 3 (Pg. 51)

##  Exercise 2
2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(dep_delay), dep_time)
```

##  Exercise 3
3. Sort flights to find the fastest flights.
```{r}
arrange(flights, air_time)
```

##  Exercise 4
4. Which flights traveled the longest? Which traveled the shortest?
```{r}
#longest and shortest time:
arrange(flights, desc(air_time))
arrange(flights, air_time)

#longest and shortest distance:
arrange(flights, desc(distance))
arrange(flights, distance)
```

# Chapter 3 (Pg. 54)

##  Exercise 2
2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, dep_time, dep_time)
```
Nothing special happens, the chosen variable is just displayed once.

##  Exercise 4
4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME"))
```
This does not surprise me because the select helpers ignore the case by default. This can be changed if you specify to ignore the case, for example:
```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```
# Chapter 3 (Pg. 58)

##  Exercise 1
1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they're not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
mutate(flights, dep_time = ((dep_time %/% 100) * 60) + (dep_time %% 100), sched_dep_time = ((sched_dep_time %/% 100) * 60) + (sched_dep_time %% 100))
```

##  Exercise 2
2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
arrange(flights, air_time)
arrange(flights, (arr_time - dep_time))
```
We would expect to see equal values, but because arr_time is represented differently than dep_time, the calculation is off. To fix this, we can make conversions similar to those in question 1.1:
```{r}
mutate(flights, dep_time = ((dep_time %/% 100) * 60) + (dep_time %% 100), sched_dep_time = ((sched_dep_time %/% 100) * 60) + (sched_dep_time %% 100), arr_time = ((arr_time %/% 100) * 60) + (arr_time %% 100), sched_arr_time = ((sched_arr_time %/% 100) * 60) + (sched_arr_time %% 100))
```

##  Exercise 3
3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

Because the time of departure should be once the number of minutes that the flight was delayed has elapsed since the scheduled departure time for any flight, dep_time = sched_dep_time + dep_delay.

##  Exercise 4
4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().
```{r}
filter(flights, min_rank(desc(dep_delay)) <= 10)
#if there is a tie, min_rank will give more than 10.
```