```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 1 (Pgs. 6-7)

##  Exercise 1
1. Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data=mpg)
```
There is just a blank space for a graph.

## Exercise 2
2. How many rows are in mtcars? How many columns?
```{r}
dim(mpg)
nrow(mpg)
ncol(mpg)
```
There are 234 rows and 11 columns in the dataset **mpg**.

##  Exercise 3
3. What does the drv variable describe? Read the help for ?mpg to find out.
```{r, eval=FALSE}
?mpg
```

```{r}
unique(mpg$drv)
```
The variable **drg** is a factor variable that takes the following values:

- "f" = front-wheel drive
- "r" = rear-wheel drive
- "4" = 4-wheel drive

##  Excercise 4
4. Make a scatterplot of hwy versus cyl.
```{r}
ggplot(data=mpg,aes(x=hwy,y=cyl)) +
  geom_point() + 
  xlab("Highway Miles Per Gallon") +
  ylab("Number of Cylinders")
```


##  Excercise 5
5. What happens if you make a scatterplot of class versus drv? Why is the plot not useful?
```{r}
ggplot(data=mpg,aes(x=class,y=drv)) + 
  geom_point() + 
  xlab("Type of Car") +
  ylab("Type of Drive")
```
Scatter plots are not meant to visualize the relationship between two categorical/qualitative variables. 

# Chapter 1 (Pg. 12)

##  Exercise 1
1. What's gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) +
geom_point(
mapping = aes(x = displ, y = hwy, color = "blue")
)
```
The points are not blue because blue is being treated like a variable instead of an aesthetic property. Color = blue must be moved outside of the mapping function to turn the color of the points to blue.

```{r}
ggplot(data=mpg) +
  geom_point(mapping=aes(x=displ,y=hwy),color="blue")
```

##  Exercise 2
Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?
```{r}
#?mpg
str(mpg)
```

**manufacturer**, **model**, **cyl**, **trans**, **drv**, **fl**, and **class** are categorical while **displ**, **year**, **cty**, and **hwy** are continuous.
```{r}
ggplot(data=mpg) +
  geom_point(mapping=aes(x=displ,y=hwy,color=cty))
ggplot(data=mpg) +
  geom_point(mapping=aes(x=displ,y=hwy,size=cty))
#This works, but shape=cty or shape=displ will not work for example because cty and displ are not categorical variables. It will say "Error: a continuous variable can not be mapped to shape".
```

##  Exercise 5
5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)
```{r}
ggplot(data=mpg) +
  geom_point(mapping=aes(x=displ,y=hwy,stroke=cty))
```

```{r,eval=F}
?geom_point
```

**stroke** is like the **shape** aesthetic and the size of the outline around the points is dependent on the magnitude of **cty**.

# Chapter 1 (Pgs. 20-21)

##  Exercise 1
1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

**geom_line** for line chart, **geom_boxplot** for boxplot, **geom_histogram** for histogram, **geom_area** for area chart.

##  Exercise 2
2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions:
```{r}
ggplot(
data = mpg,
mapping = aes(x = displ, y = hwy, color = drv)
) +
geom_point() +
geom_smooth(se = FALSE)
```

##  Exercise 3
3. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?

```{r}
ggplot(
data = mpg,
mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE
) +
geom_point(show.legend = FALSE) +
geom_smooth(se = FALSE, show.legend = FALSE)
```

show.legend = FALSE removes the legend. This makes the plot look cleaner and neater.

##  Exercise 4
4. What does the se argument to geom_smooth() do?

The `se` argument controls the standard error boundary around the smooth curve.

##  Exercise 5
5. Will these two graphs look different? Why/why not?

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. These are the same graph.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.

```{r}
#?diamonds
ggplot(data=diamonds) +
  geom_boxplot(aes(x=color,y=carat,color=color)) + 
  facet_wrap(~cut) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) + 
  guides(color=guide_legend(title="Color"))+
  xlab("") + ylab("Weight")
```


## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables.
```{r}
ggplot(data=diamonds) + 
  geom_point(aes(x=carat,y=price,alpha=clarity,color=color)) 
```
